---
title: "Introduction for the GeoRTS Package"
author: "Cristian Pachacama"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(knitr)
```



## **Overview**

The GeoRTS package provides a set of tools for reconstruction (by interpolation) of geographical time series


## **Installation**

Install the stable version from [CRAN](https://CRAN.R-project.org/package=GeoRTS):

```{r message=FALSE, warning=FALSE, eval=FALSE, echo=TRUE}
install.packages("GeoRTS")
```


or install the development version from [Github](https://github.com/InstitutoInvestigacionesEconomicasPUCE/GeoRTS):

```{r message=FALSE, warning=FALSE, eval=FALSE, echo=TRUE}
# install.packages("devtools")
devtools::install_github("InstitutoInvestigacionesEconomicasPUCE/GeoRTS")
```

## **Key Features**

### Cleaning of time series

First we need to carge the data, so we use:

```{r,eval=F}
library(GeoRTS)
library(gridExtra)
data(geoRTS)
```

Then we have in our environment five new objects: `TS` , `positions.TS` , `weights.TS` , `positions.RTS` , `weights.RTS`



```{r,echo=FALSE, message=FALSE, warning=FALSE}
library(stlplus)
library(dplyr)
library(ggplot2)
library(reshape2)
library(gridExtra)
library(leaflet)
library(highcharter)
load(file="func_borr.RData")
load(file="rts.RData")

```


Where `TS` is a multivariate "ts" object, which has missing values:

```{r}
kable(TS[10:25,])
```

Then applyinng `rts_clean` function to `TS`, by consider the parameter of seasonality 12: 

```{r,eval=T, message=FALSE, warning=FALSE}
TS_clean = rts_clean(TS,seasonality = 12)
kable(TS_clean[10:25,])
```

Using the `rts_plotClean` we can plot the original and cleaned function:

```{r, message=FALSE, warning=FALSE,fig.align="center"}
pl = list()
for (i in seq(dim(TS)[2])) {
  pl[[i]] = rts_plotClean(TS,i)
}

grid.arrange(grobs=pl,nrow=dim(TS)[2])
```


### Reconstruction of time series

To reconstruct (by IDW) unknownn data of time series associated to `positions.RTS` (latitude and longitud in meters) we use `geoRTS` function, that consider an aproximation based in convex combination of original time series in `TS`, considering the distance between `TS` given by `positions.TS`, also it could be consider some factors as `weight.TS` related with the local geographical influence of time series.


```{r}
RTS = geoRts(TS_clean,positions.TS,weights.TS,positions.RTS,weights.RTS)
kable(RTS[10:25,])
```


### Plot the RTS data

Then we could plot both series (original and reconstructed) with the function `rts_plotGroup`

```{r,fig.align="center"}
rts_plotGroup(TS_clean,RTS)
```

Finally we could represent the date in a map with function `rts_map`

```{r,fig.align="center"}
rts_map(positions.TS,positions.RTS,scale=10^4,weights.TS,weights.RTS)
```
